---
title: "hw02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Step 1: Smell test the data
Load the data set and try to figure out what the data set looks like. 
```{r}
library(gapminder)
library(tidyverse)
typeof(gapminder)
class(gapminder)
ncol(gapminder)
nrow(gapminder)
length(gapminder)
str(gapminder)
```
### Step 2: Explore individual variables
```{r}
summary(gapminder$continent) 
```

Take all countries in Europe that have a GDP per capita greater than 40000, and select all variables except gdpPercap. 

```{r}
gapminder %>%
  filter(gdpPercap>40000 & continent=="Europe") %>% 
  select(-gdpPercap) %>% 
  knitr::kable()

```


### Step 3: Explore various plot types
- **Scatter plot**

Look at the relationship between gdpPercap and lifeExp.

```{r}
gapminder %>% 
    filter(continent=="Europe") %>% 
    ggplot(aes(x=log(gdpPercap), y=lifeExp))+
    geom_point()
```

Obviously, higher gdp per capita usually leads to longer lifetime in Europe.

- **Density plot**

Show the distribution of population in each continent.
```{r}
library(tidyverse)
gapminder %>% 
    ggplot(aes(pop,fill= continent))+
    scale_x_log10()+
    geom_histogram(bins=20)
```

It can be concluded that Africa accounts for the highest pencentage of population in the world. And there is a high variance of population in Asia.

- **Line plot**

Focus on the trend of population of each city in Europe and Asia.

```{r}
gapminder %>%  
  filter(continent=="Asia"|continent=="Europe") %>% 
  ggplot(aes(year, pop, group=country, color=continent)) +
  scale_y_log10()+
  geom_line()
```

These years the countries in Asia have seen a remarkable increase on population. Whereas most of countries in Europe keep steady generally on that.

### Go further

1. Logical Operators

Get the data for Rwanda and Afghanistan.
 
```{r}
filter(gapminder, country=="Rwanda"|country=="Afghanistan") %>% 
knitr::kable()
```

"||" couldn't achieve filter here. Since it evaluates left to right examining only the first element of each vector. 

```{r}
filter(gapminder, country=="Rwanda"||country=="Afghanistan")
```

2. More dplyr: `group_by` and `transmute` 

Sum over the population of each continent in 2007. 
```{r}
library(dplyr)
a <- group_by(gapminder, continent)  
a <- filter(a, year==2007)
a <- transmute(a, cum = cumsum(as.numeric(pop))) 
filter(a, cum==max(cum) ) %>% 
  knitr::kable()
```

`group_by` takes an existing tbl and converts it into a grouped tbl.

`transmute` adds new variables and drops existing variables.
